\blindfootnote{Parts of this fictional story are drawn from a non-fictional interview on the Soft White Underbelly YouTube channel.}

\blindfootnote{The title is a weak pun: a ship is only considered to be truly at sea ``after the breakwater'', but in this story the Breakwater is a proper noun.}

I had been working with Kipper for about six months -- we were working in one of those forgotten dockyards on the south coast\smallmarginnote{Portland} -- when I decided to try my luck with his granddaughter, Charly. If that sounds like a violation of some unwritten code, what can I say to help you understand that it wasn't?

To begin with, I liked Kipper a great deal. Before I met him, I had found myself in a very bad situation. I'll spare you the details, but I had nowhere to live, no job, basically no money, and no way out. My uncle, my father's brother, gave me Kipper's number, and that was when things started to get better. He let me stay with him until I could afford to lodge elsewhere; he got me my job. Most of all, he showed me that I could earn a living doing something that I was good at: he taught me about heaving lines and dipping the eye; he showed me how to give my knife a better handle by wrapping paracord around the wood; he even tried to teach me how to splice a steel wire rope. (Although I'm ashamed to say I forgot that last lesson almost as soon as I had seen it.) To me, he saved my life as surely as if he'd dragged me out of a burning building. And he was a big man: we used to joke that the rest of us could fit a whole fist through his wedding ring, and if I still had my milk teeth I would've said he seven feet tall.

You also have to take into account that Charly was not exactly immaculate either. I don't say that with any kind of condemnation or superiority -- indeed, that was half the reason I liked her -- but that was how she was. She was nineteen, with a tongue piercing and an ongoing investigation for insurance fraud. No one was stealing anyone else's innocence. And, as I said, I liked her; I don't have to tell you that she was nice to look at, and she seemed to have every ounce of her grandfather's bullish humanity.

\divsep

I had planned to make my move one Friday night in November. One of the guys that myself and Kipper were working with in the dockyard was retiring, and we were all getting to together -- with our wives, girlfriends, etc -- at the Breakwater\smallmarginnote{Royal Breakwater Hotel}, the pub nearest the dockyard gates, next to the marina. It was one of those pleasant autumn nights, dry and clear, and cold enough to make it feel wonderful to come inside and stand by an open fire.

I decided to arrive an hour or so later than the others. I reasoned that starting a bit later, drinking-wise, not to mention an air of mystery, couldn't hurt my case with Charly. And I wasn't wrong -- at least, not in any sense I had anticipated. Once I had negotiated my way across the crowded taproom, towards the bar, and shaken hands with Kipper and the rest of the guys, I couldn't see Charly or any of her friends. But, before I could order a drink, I felt someone squeeze my shoulder, and turned round and saw her in front of me.

`Hey,' she said. I could smell the whisky on her breath -- she was going through a fireball-drinking phase -- and I liked it.

`Hey.'

`Can I talk to you?' she asked. I felt my loins stirring.

`Of course.'

She led me into one of the side rooms, which was quieter than the main bar, and we talked for a couple of minutes. It turned out that one of the girls there, Alex, was going through some kind of crisis. There didn't seem to be anything wrong with her physically -- one of the ladies present was a nurse -- except that her heart was racing a bit, and she was obviously not herself. She was as white as a sheet, and she was answering the other girls' questions only with some difficulty.

`I think she's just panicking over something,' Charly told me. `Would you -- I hate to ask this of you -- but would you take her home? If my granddad trusts you, I can't think that you would do anything to hurt her. I know she likes you, and -- don't hate me for saying this -- but I heard you had some personal experience with this sort of thing.'

`Okay,' I said. `I'll see what I can do.'

I tended to keep my car in the dockyard, where there was plenty of space; I was lodging with an older lady who lived close to the castle, so it was only a twenty minute walk into work. Alex was, as I said, in a state, and I never had chance to have drink, so I hurriedly picked up my old hatchback, and, with Charly's help, bundled the former into the front passenger seat. In the clearer illumination of my car's courtesy light, I saw a woman in her late twenties, not much more than five feet, with a vaguely Mediterranean look about her, black hair and green eyes. (Charly, I couldn't help but notice, had gone back inside without even wishing me luck.)

`Well,' I said, `where's home?'

`Across town. The other side of the causeway, I mean.' She bit her top lip thoughtfully. `But I can't go back there like this. I live with my mum. She'll think I was doing drugs or something. Could we go to yours, just for a bit?'

`Well, I never say no to a woman offering to go home with me,' I said. And I regretted the insinuation as soon as I had uttered it. `Obviously -- there's no way of saying this without it being weird -- we don't have to do anything like that.'

\divsep

As luck would have it, my landlady had gone away for the weekend, so we had the run of the house.

`Tea,' I asked, `or something stronger?'

`Tea will be fine. Thank you.' Then she remembered herself. `Milk, no sugar. Thank you.'

`Ah, a refined lady. Just like myself.'

`You're a refined lady?' she asked; and I tried to laugh heartily enough to put her at ease, but no so much as to come across as a psycho. I'm still not sure whether I pulled it off.

We stood in the kitchen for a few minutes, exchanging the usual pleasantries about the local area. Then I gestured towards the living room and its pair of armchairs. Was there something she wanted to talk about?

`Yes,' she said. `Very much so.'

Once we had sat down, she looked nervously out of the window, into the inky blackness of the night, for about a minute. Then she said, `I don't really know where to begin.'

`At the beginning?' I offered sheepishly. Then she began to speak more fluently.

`I had two mothers for the first twelve years of my life. Then they broke up, and I moved in with my dad. It was my decision to live with my dad, with my stepmother and my brothers. Something about me triggered some part of him; of all of his children, I was the only one he puts his hands on; I don't believe he ever hit my stepmother. That's sort of a thread throughout my life: with my first boyfriend, things became violent eventually -- he beat the living daylights out of me a few times; then my ex-husband -- we were together almost ten years -- he had never hit a woman before, and he ended up beating me within a year of us getting together. I'm not some like fragile victim that thinks that everything's my fault, but I do think like: three completely different men, and I'm the common denominator. There has to be something about me that, as a minimum, brings out the worst in men. The abuse from dad, it was just physical, but it was pretty bad. I remember when I was sixteen, one of my teachers noticed the bruises on my face -- I had tried to cover them up with makeup -- and I lied and said that one of those giant tin cans had fallen on me at work. I don't know why; I guess I didn't want to get him in trouble. I just didn't want him to do it again. And I guess I thought, I mean, I loved him too. And I guess I thought, if I said something, I would be taken away from him, or things would get worse. I'm not nearly as angry now as I was back then, but one thing that still screws with my head is that my son -- he's the same age now as when my dad first hit me -- did exactly the same thing as I did to get hit, and' -- she paused for a moment -- `and it was so very easy not to hit him. I was angry and disappointed, but all I had to do was look at his face, and no part of me wanted to lash out or hurt him. I don't understand why it was so easy to hit me.'

She took a deep breath and rubbed her thigh. Had it been daylight, I think she might have liked to look out of the window for a few moments. Then she continued: `Of course, I moved out as soon as I could. I managed to get myself a flat actually just before I turned eighteen, and then within a year I got pregnant with my son. By the time he was nine months old, I was working full time again, in a nursing home, and then I managed to injure myself, badly enough to get six weeks off work, and they gave me pills for the pain. After the pills ran out, I started to go out of my mind. It sounds stupid, especially since so many of my family are addicts, but I honestly didn't know anything about withdrawal. My aunt realised what was happening and gave me a much stronger pill; she said it would make me feel better. And it really did. I was lying in bed in agony, and within an hour I had cleaned the whole flat and was on the floor playing with my son. Then in four months I went from swallowing the pills to snorting them, and then from snorting them to injecting them, and then onto heroin, because it was cheaper. And from there it's a story you must have heard a million times already. I stole from everyone I loved. My son's father applied for an Emergency Residence Order and ended up being awarded full custody; I did a twelve-week residential program, completed it, then found myself shooting up two days later. At that point, I had screwed over everyone I knew, and it all pretty much caught up with me.

`Church had always been part of my life. Obviously, I wasn't going during the period that I was injecting, but it was still part of my world. So, when there was no one else that would help me, I turned to the people I knew from church. One of the flower ladies had a who granddaughter had gone through the same sort of thing that I was going through, and she said she had a sister, Cathy, that I could stay with her for as long as I needed. She lived five hours' drive away; my plan was to run away, get better and come back for my son.

`I still very much regret accepting that invitation, but I guess as with anything else in life and with anyone there's good in the bad. There was a lot of good and a lot of healing that came out of that situation. Cathy's husband, Chris, was away a lot; he worked one month on, one month off -- something to do with the rigs, I think. But they grandchildren living with them -- two boys, four and six -- and those kids just breathed life into me. Every day they would tell me they loved me. I got a job stacking shelves. Three months of that, and I started feeling like a normal person again.

`I started going to church again too. I didn't believe any more, of course, but Cathy and her family went, so I went out of respect for her. She used to make me wear these dresses.' She let out half a chuckle. `I honestly don't think I had worn a dress in my life before then. It was summertime and, right before we went to church, Chris, Cathy's husband, used to take pictures of us all in the front garden, and he used to turn the camera round and show me and tell me how good I looked. I had just turned twenty-one. After a while, I started to believe it.

`One Sunday, after lunch, Chris drove me out into the countryside to look at a cottage together. He had inherited some money, and he was thinking of buying the place and renovating it; him and Cathy had talked about letting me live there for a while, like as a caretaker. Anyway, he said, it was a nice part of the world; it would be worth showing me, even if the deal didn't go through. I remember there was a gate locked with a chain, which he locked, but the car wouldn't fit through the gap, and we had to walk the rest of the way through the woods. The path seemed to go on forever, and then I remember him on a tree branch; I think he was trying to make himself look bigger, but he didn't need to; he must have been six foot five at least. Something in my gut told me to run, but where was I supposed to run to, exactly? I just froze. He was never violent at all, just imposing; I was still wearing the dress I'd worn to church. I didn't tell him no; I didn't cry out. I just lay there and took it, like I did when I had a needle in my arm. When I was still doing heroin, I slept with some disgusting -- actually some really vile men -- for drugs or money, so I just tried to imagine I was somewhere else. I remember feeling like it was a dream, like it wasn't really happening. Then, before he could finish what he doing, he heard a car coming -- it was the landowner -- and he told me to hurry up and put my clothes back on. But I couldn't hurry. Everything was in slow motion. Then I remember the landowner showing me round the house, and I just acted like nothing had happened.

`I guess you're asking yourself why I didn't just leave that day. But it's not like in a TV show. Where would I go? Back home, everyone was furious with me, and with good reason; I was a horrible person back then. I deserved it. A few days later, he snuck into my room in the middle of the night, and I let him do what he came there to do. Nothing good would have come from fighting him, and I just kept thinking about those two boys in the next room. They had no idea what was going on, and I wanted to keep it that way. In the nine months I was staying there, that must have happened fifty or sixty times.

`In the end, I managed to save a bit of money, and enough time had passed that people back home weren't so angry with me anymore. I told Cathy and her husband that I was planning move back to the south coast, and he offered to drive me there. I remember we drove for a hundred miles or so, and then he stopped at a service station and told me that we were stating in a hotel for the night, which surprised me because he'd always bragged about how, when he working, he could stay awake for days at a time. I remember it was the week before Christmas; it was a bit cold when we got into the room. Part of me was just amazed; he must have been in his late fifties, but he had the sex drive of a seventeen-year-old boy. He didn't stop for the whole fourteen hours we were in that room together. If he did something that hurt me physically, or if I tried to wriggle away, he never gave me any kind of a break. He was such a huge man, he would just lift me up and put me back where I was. After a few hours, it got to a point where I couldn't escape any more, even in my own mind. Like I had to be present.'

I could sense what she was about to say, and it gave me a feeling in my stomach like looking over the edge of a cliff. I was mortified, almost literally. I was frozen in my seat.

`He gave me my first orgasm,' she said. `The very first orgasm of my life. That was so fucking hard to say. After that, he dropped me off, and thank God I never saw him again.'

`I want to say that it wasn't your fault,' I said. `And of course it wasn't your fault, but I also feel like there's nothing good I can say in response to what you've told me tonight.'

She just nodded. Then I asked her a question.

`The things you told me about, they must have happened five, ten years ago? Why is this all coming up now? Why tonight and why me?'

`I said just then that I hadn't seen Chris since he dropped me off, but actually I saw him tonight in the Breakwater, which I hadn't expected. Charly's grandfather, you call him Kipper, don't you? But his real name is Chris.'

I've never had much of a poker face, so I doubt I did a good job  of hiding my feelings, then or thereafter, try as I might. I texted Alex the next day, saying that, if she ever wanted to talk or just hang out, I would like that, but she never replied. And for weeks afterward all I could see in my mind's eye was Chris Kipper standing by the bar, joking and laughing, his short white beard and, behind it, his yellow teeth.
